### 2020-12-21 docs/manuscript/manuscript.Rmd
manuscript

---
title: Template for Oxford University Press papers
#date: "`r Sys.Date()`"
authors:
  - name: Authors
    email: emails
    address: institutions
abstract: |
  This is the abstract.
acknowledgements: |
  This is an acknowledgement.
keywords:
  - key
  - dictionary
  - word
#fontsize: 12pt
#spacing: halfline # could also be oneline
#classoptions:
#  - endnotes
bibliography: references.bib
output: rticles::oup_article
#header-includes:
#  - \usepackage[nomarkers,tablesfirst]{endfloat} # For figures and tables at end
#  - \usepackage{lineno} # For line numbering
#  - \linenumbers # For line numbering
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # By default, hide code; set to TRUE to see code
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on pages separate from text
knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment

## Load packages
library("ggplot2")
library("ggpubr")
library("drake")
```

```{r load-data}
simulation_parameters = readd("simulation_parameters")
rt_estimates_params_acc = readd("rt_estimates_params_acc")
rt_estimates_params_inacc = readd("rt_estimates_params_inacc")

error_median = with(rt_estimates_params_acc, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
error_quantile_95 = with(rt_estimates_params_acc, quantile(abs(rt.simple_mean_error), 0.95, na.rm = T)) %>% signif(digits = 2)
error_quantile_05 = with(rt_estimates_params_acc, quantile(abs(rt.simple_mean_error), 0.05, na.rm = T)) %>% signif(digits = 2)

```

# Introduction

The effective reproduction number ($R_t$) is a fundamental measure in infectious disease epidemiology. It can be defined as the average number of secondary caused by an infected individual. The $R_t$ changes over the course of an epidemic, because it is not only determined by characteristics of the pathogen but also by human behavior (e.g. face mask usage, physical distancing, and travel restrictions). If the $R_t$ is above 1, it indicates that epidemic is growing exponentially. If, by contrast, the $R_t$ drop below 1, it indicates that the number of new cases is decreasing. Furthermore, the $R_t$ can vary between geographic locations, as social behaviors vary between these locations.

From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region. However, this communication may be hindered by the fact that many of common methods for estimating $R_t$ require advanced mathematics to understand [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007]. Here, we describe how $R_t$ can be approximated by a ratio of new infections on day $t$ and day $t-d$, where d is the mean generation interval of the infection ("generation interval" is defined as the average time from when a primary case becomes infectious to when a secondary becomes infectious). The major advantage of this approximation is that its simplicity makes it easily understandable and verifiable.

## A Simple Ratio for Estimating Rt

The total number of new infections at time, $I_t$, can be calculated by summing the expected number of new cases created at $t$ by each person infected prior to $t$. The expected of number of new cases created by an infected person can be considered as a function, $n(a)$, of the age ($a$) of their infection [@wallinga_how_2007].

$$I_t = \int_a^\infty{I_{t-a}n(a)}da$$

The function $n(a)$ can be normalized to by the total number of infection an infected person is expected cause (i.e. $R$) to produce the generation interval distribution:

$$g(a) = \frac{n(a)}{R}$$

This can be substituted into equation 1 to produce:

$$\frac{I_t}{R} = \int_a^\infty{I_{t-a}g(a)}da$$
$${R} = \frac{I_t}{\int{I_{t-a}g(a)}da}$$

If we make the assumption that there is minimal variance in the generation interval distribution (i.e. $g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then equation 4 can be simplified to:

$${R} \approx \frac{I_t}{I_{t-d}}$$

We can make this estimation of $R_t$ robust to variations in the generation interval and reporting of the infection by calculating $R_t$ as the ratio of the sums of new cases within two time windows size $\tau$ days (with the center of the time periods being on days $t$ and $t-d$):

$$ R_t \approx \frac{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_i}{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_{i-d}} = \frac{n_{t}}{n_{t-d}}$$
A confidence interval can be generated for this ratio by carrying forward a confidence interval in the generation time.

# Materials and methods

# Results

## Accuracy of the simple ratio estimation of $R_t$

We assessed how accurately the simple ratio can estimate $R_t$ compared to a previously developed method by Cori et al [@cori_new_2013], which is a popular method in the field - we will refer to this method as the "Cori method". $R_t$ calculation with either the simple ratio and Cori method requires knowing the mean generation interval of the infection; however, the Cori method also requires knowing the standard deviation of the generation interval or having an empiric distribution of the generation interval.

To assess accuracy of $R_t$ estimations, we used Susceptible-Exposed-Infectious-Resolved (SEIR) model to simulate epidemics and then compare how accurately the two methods can estimate the $R_t$ values in the epidemics. We simulated `r simulation_parameters$Nsims` epidemics in a population of `r addUnits(N)` individuals. We varied the lengths of latent and infectiousness periods with each simulation. In each simulation, the population underwent changes in the $R_t$ value representing changes in population behavior (e.g. "lockdown" period or lifting of a lockdown).  **Figure \@ref(fig:sim-results) A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic. A full visualization of the all the simulations can be seen in **Figure S1**.

Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was  `r error_median` (90% range of `r error_quantile_05` to `r error_quantile_95`) (**Figure \@ref(fig:sim-results) B**). The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (**Figure \@ref(fig:sim-results) C**).

Often however, the the mean and standard deviation of the generation interval is not well known. Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer. Under these conditions, the simple ratio actually performs slightly better than than the Cori et al method (**Figure \@ref(fig:sim-results) D-E**). By requiring fewer parameters, the simple ratio is robust to mis-specification of the standard deviation. This is a legitimate concern given that estimates for the standard deviation of the generation interval for the COVID-19 have ranged from 0.95 to 4.75 days [@knight_estimating_2020].

## Simple ratio estimation of $R_t$ in COVID data

We next compared $R_t$ estimates from the Simple Ratio and Cori et al for using data from the current COVID-19 pandemic. We estimated $R_t$ in four nations using a generation interval with mean of `r mean_gi` and standard deviation of `r sd_gi`[@ganyani_estimating_2020]. **Figure \@ref(fig:real-world-data) A** shows the results for the month of October 2020. The two estimates have very good concordance, with an average difference in estimation of `r round(mean_diff,2)` (**Figure \@ref(fig:real-world-data) B**).

```{r real-world-data, fig.cap = "\\textbf{The Simple Ratio and Cori Method estimate similar Rt values for the COVID-19 pandemic}"}

```

# Discussion

# References

### 2020-12-24 code/COVID19_data_analysis.R
download covid19 data

@@ -1,0 +1,13 @@
+
+download_covid19_data = function() {
+
+  national_data = get_national_data(totals = F, source = "WHO")
+
+  national_data = national_data %>%
+    mutate(time = date) %>% select(-date) %>%
+    filter(time > as.Date("2020-04-01") & time < as.Date("2020-12-01"))
+
+  return(national_data)
+}
+
+

### 2020-12-24 code/create_figures.R
scripts to create figures

@@ -1,0 +1,64 @@
+# This R script is for making the figures for the manuscript
+# color-blind friendly palette
+cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")
+
+custom_theme = function() {
+  theme_bw() +
+    theme(legend.position = "bottom",
+          axis.title = element_text(size = 8))
+}
+
+
+create_rt_estimation_error_line <- function(data) {
+  data %>%
+    pivot_longer(cols = starts_with("error"), names_to = "method", names_prefix = "error.", values_to = "error") %>%
+    ggplot() +
+    geom_line(aes(time, error, group = paste(simulation_run, method, sep="_"), color = method), alpha = 0.2) +
+    ylab("Predicted Rt - Actual Rt") +
+    custom_theme() +
+    scale_color_manual(values=cbPalette) +
+    scale_fill_manual(values=cbPalette)
+}
+
+create_rt_estimation_error_boxplot <- function(data) {
+  data %>%
+    pivot_longer(., cols = starts_with("rt."), names_to = c("Method", "parameter"),
+                 names_pattern = "rt.([[:alpha:]]+)_([[:print:]]+)", values_to = "value") %>%
+    pivot_wider(names_from = "parameter", values_from = "value") %>%
+    mutate(error = abs(mean_error)) %>%
+    ggplot(., aes(Method, error)) +
+    geom_violin(aes(fill = Method)) +
+    geom_boxplot(width = 0.1, outlier.alpha = 0.5) +
+    ylab("log10(Estimation Error)") +
+    custom_theme() +
+    scale_y_log10(labels = scales::label_number(accuracy = 0.00001)) +
+    scale_color_manual(values=cbPalette) +
+    scale_fill_manual(values=cbPalette)
+}
+
+create_rt_estimation_error_by_rt <- function(data) {
+  data %>%
+    ggplot() +
+    geom_point(aes(r_eff, rt.simple_mean_error), size=0.5, alpha=0.3, color = cbPalette[2]) +
+    labs(y = "Estimation Error", x = "True Rt") +
+    custom_theme() +
+    scale_color_manual(values=cbPalette) +
+    scale_fill_manual(values=cbPalette)
+}
+
+create_estimate_plot <- function(data) {
+  data = data %>%
+    pivot_longer(., cols = starts_with("rt."), names_to = c("Method", "parameter"),
[…]

### 2020-12-24 code/packages.R
additional packages used

@@ -4,3 +4,6 @@ library("drake")
 library("EpiEstim")
 library("deSolve")
 library("RcppRoll")
+library("ggpubr")
+library("ggplot2")
+library("covidregionaldata")

### 2020-12-24 code/functions.R
created min_newCases and min_simLengths variables

@@ -130,11 +130,12 @@ model_pandemic <- function(initial_state_values, mean_gi, gamma_val, r0, period_
   return(output)
 }
 
-model_stochastic_parameters <- function(initial_state_values, mean_gi_range, r0_range, num_periods, period_lengths_range, transition_lengths_range) {
+model_stochastic_parameters <- function(initial_state_values, mean_gi_range, r0_range, num_periods, period_lengths_range, transition_lengths_range,
+                                        min_simLength, min_newCases) {
 
   i = 0
   sim_data = data.frame()
-  while(nrow(sim_data) < 150 | any(sim_data$new_cases < 20, na.rm = T)) {
+  while(nrow(sim_data) < min_simLength | any(sim_data$new_cases < min_newCases, na.rm = T)) {
     # i = i + 1
     # print(paste("trial...",i))
 
@@ -178,7 +179,9 @@ model_stochastic_parameters_multiwrapper <- function(simulation_parameters) {
                                                r0_range = seq(r0_range[1], r0_range[2], r0_range[3]),
                                                num_periods = num_periods,
                                                period_lengths = seq(period_lengths[1], period_lengths[2], period_lengths[3]),
-                                               transition_lengths =  seq(transition_lengths[1], transition_lengths[2], transition_lengths[3])), simplify = F)
+                                               transition_lengths =  seq(transition_lengths[1], transition_lengths[2], transition_lengths[3]),
+                                               min_simLength = min_simLength,
+                                               min_newCases = min_newCases), simplify = F)
     )
 }
 

### 2020-12-24 code/functions.R
removed 0.1 correction in simple method rt estimation

@@ -195,7 +195,7 @@ estimates_rt.simple <- function(time, Is, mean_gi, tau) {
   df = df %>%
     mutate(new_cases_sum = roll_sum(Is,  tau, align="center", fill = c(NA, NA, NA))) %>%
     mutate(new_cases_sum_gi_days_ago = lag(new_cases_sum, mean_gi)) %>%
-    mutate(rt.simple_mean = (new_cases_sum+0.1)/(new_cases_sum_gi_days_ago+0.1))
+    mutate(rt.simple_mean = (new_cases_sum)/(new_cases_sum_gi_days_ago))
 
   # Confidence Interval.using beta representation of Clopper-Pearson method
   # (instead of using poisson.test(), which uses binom.test indirectly)

### 2020-12-24 code/plan.R
plan name

@@ -1,4 +1,4 @@
-run_simulation <- drake_plan(
+plan <- drake_plan(
   # Load simulation parameters
   simulation_parameters = list(
     mean_gi_range = c(4,12,1),

### 2020-12-24 code/plan.R
min_simLength and min_newCases variables

@@ -10,7 +10,9 @@ plan <- drake_plan(
     initial_state_values = c(S = 2e7-100,
                               E = 0,
                               I = 1000,
-                              R = 0)
+                              R = 0),
+    min_simLength = 150,
+    min_newCases = 20
   ),
 
   # Run Epidemic Simulations

### 2020-12-24 code/plan.R
rt_estimates var name change

@@ -19,10 +19,10 @@ plan <- drake_plan(
   epidemic_simulations = model_stochastic_parameters_multiwrapper(simulation_parameters),
 
   # Estimate Rt from simulation with accurate parameters
-  rt_estimates_params_acc = map(epidemic_simulations, function(i) estimates_rt(i,
-                                                    mean_gi = attributes(i)$metadata["mean_gi"],
-                                                     gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
-                                                     error.mean = 0, error.sd = 0)) %>% bind_rows(., .id = "simulation_run"),
+  rt_estimates = map(epidemic_simulations, function(i) estimates_rt(i,
+                      mean_gi = attributes(i)$metadata["mean_gi"],
+                       gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
+                       error.mean = 0, error.sd = 0)) %>% bind_rows(., .id = "simulation_run"),
 
   # Estimate Rt from simulation with inaccurate parameters
   rt_estimates_params_inacc = map(epidemic_simulations, function(i) estimates_rt(i,

### 2020-12-24 code/plan.R
analysis plan for covid19 data

@@ -29,4 +29,15 @@ plan <- drake_plan(
                                    mean_gi = attributes(i)$metadata["mean_gi"],
                                    gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
                                    error.mean = 0, error.sd = 2)) %>% bind_rows(., .id = "simulation_run")
+  # Real-world data
+  COVID19_params = list(mean_gi = 4.83, sd_gi = 1.72),
+  national_data_raw = download_covid19_data(),
+  national_data = national_data_raw %>%
+    group_by(country) %>% nest() %>%
+    mutate(rt.simple = map(data, function(x) estimates_rt.simple(time = x$time, Is = x$cases_new, mean_gi = round(COVID19_params$mean_gi), tau = 7))) %>%
+    mutate(rt.cori = map(data, function(x) estimates_rt.cori(time = x$time, Is = x$cases_new, mean_gi = COVID19_params$mean_gi, sd_gi = COVID19_params$sd_gi, tau = 7))) %>%
+    mutate(data = map2(data, rt.simple, full_join, by = "time")) %>% select(-rt.simple) %>%
+    mutate(data = map2(data, rt.cori, full_join, by = "time")) %>% select(-rt.cori) %>%
+    unnest(data) %>%
+    mutate(estimation_diff = abs(rt.cori_mean - rt.simple_mean)),
 )

### 2020-12-24 code/plan.R
split rt_estimates_params_inacc into rt_estimates_meanerr and rt_estimates_sderr in plan

@@ -24,11 +24,18 @@ plan <- drake_plan(
                        gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
                        error.mean = 0, error.sd = 0)) %>% bind_rows(., .id = "simulation_run"),
 
-  # Estimate Rt from simulation with inaccurate parameters
-  rt_estimates_params_inacc = map(epidemic_simulations, function(i) estimates_rt(i,
-                                   mean_gi = attributes(i)$metadata["mean_gi"],
-                                   gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
-                                   error.mean = 0, error.sd = 2)) %>% bind_rows(., .id = "simulation_run")
+  # Estimate Rt from simulation with inaccurate mean
+  rt_estimates_meanerr = map(epidemic_simulations, function(i) estimates_rt(i,
+                               mean_gi = attributes(i)$metadata["mean_gi"],
+                               gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
+                               error.mean = 2, error.sd = 0)) %>% bind_rows(., .id = "simulation_run"),
+
+  # Estimate Rt from simulation with inaccurate sd
+  rt_estimates_sderr = map(epidemic_simulations, function(i) estimates_rt(i,
+                           mean_gi = attributes(i)$metadata["mean_gi"],
+                           gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
+                           error.mean = 0, error.sd = 2)) %>% bind_rows(., .id = "simulation_run"),
+
   # Real-world data
   COVID19_params = list(mean_gi = 4.83, sd_gi = 1.72),
   national_data_raw = download_covid19_data(),

### 2020-12-24 _drake.R
add additional source functions

@@ -1,8 +1,10 @@
 source("./code/packages.R")
 source("./code/functions.R")
 source("./code/plan.R")
+source("./code/create_figures.R")
+source("./code/COVID19_data_analysis.R")
 
 drake_config(
-  run_simulation,
+  plan,
   verbose = 2
 )

### 2020-12-24 docs/manuscript/manuscript.Rmd
manuscript update

@@
 ---
-title: Template for Oxford University Press papers
-#date: "`r Sys.Date()`"
+title: "A Simple Ratio for Estimating the Effective Reproduction Value of an Epidemic"
 authors:
@@
 abstract: |
-  This is the abstract.
+  ABSTRACT
 acknowledgements: |
-  This is an acknowledgement.
+  Acknowledgements
+output:
+  bookdown::pdf_book:
+    base_format: rticles::oup_article
+    number_sections: no
+    extra_dependencies: float
+    keep_tex: yes
+    includes:
+      in_header: header.tex
+  bookdown::word_document2:
+    number_sections: no
+    extra_dependencies: float
+    keep_tex: yes
+    includes:
+      in_header: header.tex
+bibliography:
+- packages.bib
+- references.bib
 keywords:
@@
 - word
-#fontsize: 12pt
-#spacing: halfline # could also be oneline
-#classoptions:
-#  - endnotes
-bibliography: references.bib
-output: rticles::oup_article
-#header-includes:
-#  - \usepackage[nomarkers,tablesfirst]{endfloat} # For figures and tables at end
-#  - \usepackage{lineno} # For line numbering
-#  - \linenumbers # For line numbering
 ---
@@
 knitr::opts_chunk$set(fig.env="figure") # Latex figure environment
+knitr::opts_chunk$set(eval.after = "fig.cap")
+knitr::opts_chunk$set(echo = F,
+                      message = F, 
+                      warning = F)
 
 ## Load packages
+library("tidyverse")
 library("ggplot2")
@@
 library("drake")
+library("english")
+library("bibtex")
+library("here")
+
+# Set working directory
+here::i_am("docs/manuscript/manuscript.Rmd")
+
+# Get Figure creation functions
+source(here::here("code", "create_figures.R"))
+
+# Convert number to text 
+# Ref: https://5harad.com/mse125/r/visualization_code.html
+addUnits <- function(n) {
+    labels <- ifelse(n < 1000, n,  # less than thousands
+                     ifelse(n < 1e6, paste0(round(n/1e3), ' thousand'),  # in thousands
+                            ifelse(n < 1e9, paste0(round(n/1e6), ' million'),  # in millions
+                                   ifelse(n < 1e12, paste0(round(n/1e9), ' billion'), # in billions
+                                          'too big!'
+                                   ))))
+    return(labels)
+}
+
+# Covert a vector to a english list
+vector_to_englishList <- function(x) {
+  last_item = tail(x, 1)
+
+  englishList = paste(head(x, -1), collapse = ", ")
+  englishList = paste(englishList, last_item, sep = ", and ")
+
+  return(englishList)
+}
+
+# Function between to numbers
 ```
@@
 simulation_parameters = readd("simulation_parameters")
-rt_estimates_params_acc = readd("rt_estimates_params_acc")
-rt_estimates_params_inacc = readd("rt_estimates_params_inacc")
+rt_estimates = readd("rt_estimates")
+national_data = readd("national_data")
+COVID19_params = readd("COVID19_params")
 
-error_median = with(rt_estimates_params_acc, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
-error_quantile_95 = with(rt_estimates_params_acc, quantile(abs(rt.simple_mean_error), 0.95, na.rm = T)) %>% signif(digits = 2)
-error_quantile_05 = with(rt_estimates_params_acc, quantile(abs(rt.simple_mean_error), 0.05, na.rm = T)) %>% signif(digits = 2)
+N = simulation_parameters$initial_state_values %>% sum()
+error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
+error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
+error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
 
+national_data_filtered = national_data %>% 
+  filter(time > as.Date("2020-04-01") & time < as.Date("2020-12-01")) %>% 
+  filter(cases_new > 19)
+
+estimation_diff = national_data_filtered$estimation_diff
+estimation_diff_median = estimation_diff %>%  median(., na.rm=T) %>% signif(digits = 2)
+estimation_diff_quantile_975 = estimation_diff %>% quantile(., 0.975, na.rm = T) %>% signif(digits = 2) %>% format(scientific = F)
+estimation_diff_quantile_025 = estimation_diff %>% quantile(., 0.025, na.rm = T) %>% signif(digits = 2) %>% format(scientific = F)
+
+# Get packages used for this project
+packages <- readLines(here::here("code", "packages.R")) %>% #read file
+              sub('[^\"]+\"([^\"]+).*', '\\1', .)    
+package_citation_ids =  read.bib(here("./docs/manuscript", "packages.bib")) %>% 
+  names(.) %>% regmatches(., m = regexpr(pattern = "^R-.+", text = .)) %>% 
+  paste("@", ., sep="") %>% paste(collapse = "; ")
 ```
 
-# Introduction
+# Introduction {-}
 
@@
 From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region.
-However, this communication may be hindered by the fact that many of common methods for estimating $R_t$ require advanced mathematics to understand [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
+However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
 Here, we describe how $R_t$ can be approximated by a ratio of new infections on day $t$ and day $t-d$, where d is the mean generation interval of the infection ("generation interval" is defined as the average time from when a primary case becomes infectious to when a secondary becomes infectious).
+
+$$R_t \approx \frac{I_t}{I_{t-d}}$$
+
 The major advantage of this approximation is that its simplicity makes it easily understandable and verifiable.
+Anybody with basic math skills and access to the new case counts can calculate this ratio to estimate the $R_t$.
 
-## A Simple Ratio for Estimating Rt
+## A Simple Ratio for Estimating Rt {-}
 
@@
 
-$$I_t = \int_a^\infty{I_{t-a}n(a)}da$$
+\begin{equation} 
+  I_t = \int_a^\infty{I_{t-a}n(a)}da
+  (\#eq:eq2)
+\end{equation}
 
@@
 
-$$g(a) = \frac{n(a)}{R}$$
+\begin{equation} 
+  g(a) = \frac{n(a)}{R_t}
+  (\#eq:eq2)
+\end{equation}
 
-This can be substituted into equation 1 to produce:
+Equation (2) can be substituted into Equation (1) to produce:
 
-$$\frac{I_t}{R} = \int_a^\infty{I_{t-a}g(a)}da$$
-$${R} = \frac{I_t}{\int{I_{t-a}g(a)}da}$$
+\begin{equation} 
+  \frac{I_t}{R_t} = \int_a^\infty{I_{t-a}g(a)}da
+  (\#eq:eq3)
+\end{equation}
+
+\begin{equation} 
+  R_t = \frac{I_t}{\int{I_{t-a}g(a)}da}
+  (\#eq:eq4)
+\end{equation}
 
 If we make the assumption that there is minimal variance in the generation interval distribution (i.e.
-$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then equation 4 can be simplified to:
+$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then equation (4) can be simplified to:
 
-$${R} \approx \frac{I_t}{I_{t-d}}$$
+\begin{equation}
+  R_t \approx \frac{I_t}{I_{t-d}}
+  (\#eq:simple-ratio-eq)
+\end{equation}
 
@@
 
-$$ R_t \approx \frac{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_i}{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_{i-d}} = \frac{n_{t}}{n_{t-d}}$$
-A confidence interval can be generated for this ratio by carrying forward a confidence interval in the generation time.
+\begin{equation} 
+  R_t \approx \frac{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_i}{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_{i-d}} = \frac{n_{t}}{n_{t-d}}
+  (\#eq:eq6)
+\end{equation}
 
-# Materials and methods
+A confidence interval can be generated for this ratio by carrying forward a confidence interval in the generation interval.
 
-# Results
+# Materials and methods {-}
 
-## Accuracy of the simple ratio estimation of $R_t$
+## Simulation {-}
 
-We assessed how accurately the simple ratio can estimate $R_t$ compared to a previously developed method by Cori et al [@cori_new_2013], which is a popular method in the field - we will refer to this method as the "Cori method".
+A deterministic Susceptible-Exposed-Infectious-Resolved (SEIR) model was used to create epidemic simulations.
+No birth rate, death rate, or R to S transition rates were included.
+We included `r simulation_parameters$Nsims` simulated epidemics in our analysis.
+For each epidemic, the mean generation interval of the infection was randomly selected from `r simulation_parameters$mean_gi_range[1]` to `r simulation_parameters$mean_gi_range[2]` days.
+The generation interval was randomly divided into latent and infectious periods.
+Each simulation was divided in to `r as.english(simulation_parameters$num_periods[1])` `r simulation_parameters$period_lengths[1]` day periods with differing $R_0$ values (representing changes in social behaviors in the populations); and the $R_0$ in these periods was randomly selected from `r simulation_parameters$r0_range[1]` to `r simulation_parameters$r0_range[2]`.
+Between each `r simulation_parameters$period_lengths[1]` day period, a transition period was included in which the $R_0$ would gradually change following a logistic curve.
+The length of the transition period was randomly selected from `r simulation_parameters$transition_lengths[1]` to `r simulation_parameters$transition_lengths[2]` days.
+Only simulations that lasted at least `r simulation_parameters$min_simLength` and never had less than `r simulation_parameters$min_newCases` cases per day were included in final analysis.
+
+## Rt Esimation {-}
+
+Custom R code was used to generate the simple ratio $R_t$ estimations, and the EpiEstim R package was used to generate the Cori method $R_t$ estimations.
+See below for information on code availability.
+When assessing the accuracy of both methods, estimations made prior to the 15th day of each simulation was not included because both methods, as implemented here, are significantly less accurate during this early period of the epidemic.
+
+## COVID-19 Data {-}
+
+Data on daily COVID-19 cases were acquired from the covidregionaldata R package [@R-covidregionaldata]
+
+## Code and Data Availability {-}
+
+All simulations, data processing, analysis, and graphics were all created using the R statistical software using the following packages: `r vector_to_englishList(packages)` [`r package_citation_ids`].
+All code and data used to develop this project is available at [https://github.com/aryazand/simple_rt](https://github.com/aryazand/simple_rt)
+
+# Results {-}
+
+## Accuracy of the simple ratio estimation of $R_t$ {-}
+
+We assessed how accurately the simple ratio can estimate $R_t$ compared to a previously developed method by Cori et al [-@cori_new_2013], which is a popular method in the field - we will refer to this method as the "Cori method".
 $R_t$ calculation with either the simple ratio and Cori method requires knowing the mean generation interval of the infection; however, the Cori method also requires knowing the standard deviation of the generation interval or having an empiric distribution of the generation interval.
@@
 **Figure \@ref(fig:sim-results) A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
-A full visualization of the all the simulations can be seen in **Figure S1**.
+A full visualization of the all the simulations can be seen in (**Figure \@ref(fig:all-simulations)**).
 
-Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was  `r error_median` (90% range of `r error_quantile_05` to `r error_quantile_95`) (**Figure \@ref(fig:sim-results) B**).
+Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was  `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (**Figure \@ref(fig:sim-results) B**).
 The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (**Figure \@ref(fig:sim-results) C**).
+This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].
 
-Often however, the the mean and standard deviation of the generation interval is not well known.
+(ref:sim-results-cap) \textbf{The Simple Ratio approximates the effective reproduction in simualted epidemics.} One hundred epidemics were simulated using an SEIR model and the simple ratio and Cori method were used to estimate the known $R_t$ over time for each simulated epidemic.
+\textbf{(A)} The true $R_t$ (black line) and estimated $R_t$ (colored lines) over time for one of simulated epidemics.
+\textbf{(B)} The distribution of the absolute errors of both estimation methods across all simulated epidemics.
+\textbf{(C)} The estimation error of the simple ratio method is associated with size of true $R_t$
+
+```{r sim-results, fig.cap = '(ref:sim-results-cap)'}
+
+panel_A <- rt_estimates %>%
+  filter(simulation_run == 66) %>%
+  create_estimate_plot()
+
+panel_B <- rt_estimates %>%
+  filter(time > 14) %>%
+  create_rt_estimation_error_boxplot()
+
+panel_C <- rt_estimates %>%
+  filter(time > 14) %>%
+  create_rt_estimation_error_by_rt()
+
+ggarrange(panel_A, panel_B, panel_C, ncol=3, nrow=2, align = "hv", 
+          labels=c("A", "B", "C", "D", "E", "F"), common.legend = T)
+```
+
+Often however, the the mean and standard deviation of the generation interval is typically not known exactly.
+For instance, for the COVID-19 the mean generation interval estimates, with 95% credible intervals, include 3.95 (3.01 - 4.91) and 5.20 (3.78 - 6.78); and estimates of the standard deviation of the generation interval include 1.51 (0.74 - 2.97) and 1.72 (0.91 - 3.93) [@ganyani_estimating_2020].
 Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer.
-Under these conditions, the simple ratio actually performs slightly better than than the Cori et al method (**Figure \@ref(fig:sim-results) D-E**).
-By requiring fewer parameters, the simple ratio is robust to mis-specification of the standard deviation.
-This is a legitimate concern given that estimates for the standard deviation of the generation interval for the COVID-19 have ranged from 0.95 to 4.75 days [@knight_estimating_2020].
+Under these conditions, the simple ratio actually performs slightly better than than the Cori method (**Figure \@ref(fig:estimation-inaccParams)**).
 
-## Simple ratio estimation of $R_t$ in COVID data
+```{r real-world-data, fig.cap = "\\textbf{The Simple Ratio and Cori Method estimate similar Rt values for the COVID-19 pandemic}.
+\\textbf{(A)} Rt esimations for four sample nations \\textbf{(B)} Histogram displaying the absolute difference in Rt esimation across for each day of the pandemic (Apr 01 to Dec 12) for all 233 individual nations."}
+
+countries = c("United States", "Canada", "United Kingdom", "Germany")
+
+panel_A <- national_data %>%
+  filter(country %in% countries) %>%
+  filter(time %in% seq(as.Date("2020-10-01"), as.Date("2020-10-31"), by = "day")) %>%
+  pivot_longer(., cols = starts_with("rt."), names_to = c("Method", "parameter"), 
+              names_pattern = "rt.([[:alpha:]]+)_([[:print:]]+)", values_to = "value") %>%
+  pivot_wider(names_from = "parameter", values_from = "value") %>%
+  ggplot() + 
+    geom_line(aes(time, mean, group = Method, color = Method), size = 1) +
+    geom_ribbon(aes(time, ymin = quantile_025, ymax = quantile_975, group = Method, fill = Method), alpha = 0.2) + 
+    geom_hline(yintercept = 1, linetype = 2) + 
+    facet_wrap(~country) + 
+    ylab("Rt") + 
+    custom_theme() + 
+    theme(strip.background = element_blank(),
+          legend.position = "top") +
+    scale_color_manual(values=cbPalette)
+
+panel_B <- national_data_filtered %>%
+  ggplot() + 
+    geom_histogram(aes(estimation_diff)) + scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) + 
+    labs(x = "absolute difference in Rt estimation") +
+    custom_theme()
+
+ggarrange(panel_A, ggarrange(panel_B, ncol=2), nrow = 2, heights = c(2,1), labels = c("A","B"))
+```
+
+## Using the Simple Ratio to Estimate $R_t$ with real-world COVID-19 data {-}
 
 We next compared $R_t$ estimates from the Simple Ratio and Cori et al for using data from the current COVID-19 pandemic.
-We estimated $R_t$ in four nations using a generation interval with mean of `r mean_gi` and standard deviation of `r sd_gi`[@ganyani_estimating_2020].
+We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020].
 **Figure \@ref(fig:real-world-data) A** shows the results for the month of October 2020.
-The two estimates have very good concordance, with an average difference in estimation of `r round(mean_diff,2)` (**Figure \@ref(fig:real-world-data) B**).
+Visual inspection shows that the two methods of estimating $R_t$ have very good concordance.
+Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (**Figure \@ref(fig:real-world-data) B**).
 
-```{r real-world-data, fig.cap = "\\textbf{The Simple Ratio and Cori Method estimate similar Rt values for the COVID-19 pandemic}"}
+# Discussion {-}
 
+# References {-}
+
+<div id="refs"></div>
+
+# Supplementary Material {-}
+\beginsupplement
+
+\newpage
+\blandscape
+```{r all-simulations, fig.width = 11, fig.height = 6, fig.cap = "All Simulations"}
+
+rt_estimates %>%
+  create_estimate_plot() + 
+  facet_wrap(~simulation_run) + 
+  theme(strip.background = element_blank(), strip.text = element_blank())
 ```
+\elandscape
 
-# Discussion
+\newpage
+```{r estimation-inaccParams,  fig.cap = "(A) estimation with gi mean mis-specified. (B) estimation with gi sd mis-specified."}
 
-# References
+rt_estimates_meanerr <- readd("rt_estimates_meanerr")
+rt_estimates_sderr <- readd("rt_estimates_sderr")
+
+panel_A <- rt_estimates_meanerr %>% 
+  create_rt_estimation_error_boxplot()
+
+panel_B <- rt_estimates_sderr %>% 
+  create_rt_estimation_error_boxplot()
+
+ggarrange(panel_A, panel_B, ncol=2, align = "hv", labels=c("A", "B"), common.legend = T)
+```

### 2021-01-16 docs/manuscript/manuscript.Rmd
manuscript writing

@@
   labels <- ifelse(n < 1000, n,  # less than thousands
-                     ifelse(n < 1e6, paste0(round(n/1e3), ' thousand'),  # in thousands
-                            ifelse(n < 1e9, paste0(round(n/1e6), ' million'),  # in millions
-                                   ifelse(n < 1e12, paste0(round(n/1e9), ' billion'), # in billions
-                                          'too big!'
-                                   ))))
+                   ifelse(
+                     n < 1e6,
+                     paste0(round(n / 1e3), ' thousand'),
+                     # in thousands
+                     ifelse(
+                       n < 1e9,
+                       paste0(round(n / 1e6), ' million'),
+                       # in millions
+                       ifelse(n < 1e12, paste0(round(n / 1e9), ' billion'), # in billions
+                              'too big!')
+                     )
+                   ))
   return(labels)
@@
 
-  englishList = paste(head(x, -1), collapse = ", ")
+  englishList = paste(head(x,-1), collapse = ", ")
   englishList = paste(englishList, last_item, sep = ", and ")
@@
 
-N = simulation_parameters$initial_state_values %>% sum()
-error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
-error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
-error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
-
+# Comparison with COVID-19 data
 national_data_filtered = national_data %>% 
@@
 
-# Introduction {-}
+# Introduction {.unnumbered}
 
 The effective reproduction number ($R_t$) is a fundamental measure in infectious disease epidemiology.
-It can be defined as the average number of secondary caused by an infected individual.
+It can be defined as the average number of secondary cases caused by an infected individual.
 The $R_t$ changes over the course of an epidemic, because it is not only determined by characteristics of the pathogen but also by human behavior (e.g.
 face mask usage, physical distancing, and travel restrictions).
-If the $R_t$ is above 1, it indicates that epidemic is growing exponentially.
+If the $R_t$ is above 1, it indicates that the epidemic is growing exponentially.
 If, by contrast, the $R_t$ drop below 1, it indicates that the number of new cases is decreasing.
-Furthermore, the $R_t$ can vary between geographic locations, as social behaviors vary between these locations.
+Furthermore, the $R_t$ can vary between geographic locations, as social behaviors and environmental factors vary between these locations.
 
 From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region.
-However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
-Here, we describe how $R_t$ can be approximated by a ratio of new infections on day $t$ and day $t-d$, where d is the mean generation interval of the infection ("generation interval" is defined as the average time from when a primary case becomes infectious to when a secondary becomes infectious).
+However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
+Here, we describe how $R_t$ can be approximated by a ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (Equation 1).
+The average *generation interval* is defined as the average time from when a primary case becomes infectious to when a secondary case becomes infectious.
 
-$$R_t \approx \frac{I_t}{I_{t-d}}$$
-
-The major advantage of this approximation is that its simplicity makes it easily understandable and verifiable.
-Anybody with basic math skills and access to the new case counts can calculate this ratio to estimate the $R_t$.
-
-## A Simple Ratio for Estimating Rt {-}
-
-The total number of new infections at time, $I_t$, can be calculated by summing the expected number of new cases created at $t$ by each person infected prior to $t$.
-The expected of number of new cases created by an infected person can be considered as a function, $n(a)$, of the age ($a$) of their infection [@wallinga_how_2007].
-
-\begin{equation} 
-  I_t = \int_a^\infty{I_{t-a}n(a)}da
-  (\#eq:eq2)
-\end{equation}
-
-The function $n(a)$ can be normalized to by the total number of infection an infected person is expected cause (i.e. $R$) to produce the generation interval distribution:
-
-\begin{equation} 
-  g(a) = \frac{n(a)}{R_t}
-  (\#eq:eq2)
-\end{equation}
-
-Equation (2) can be substituted into Equation (1) to produce:
-
-\begin{equation} 
-  \frac{I_t}{R_t} = \int_a^\infty{I_{t-a}g(a)}da
-  (\#eq:eq3)
-\end{equation}
-
-\begin{equation} 
-  R_t = \frac{I_t}{\int{I_{t-a}g(a)}da}
-  (\#eq:eq4)
-\end{equation}
-
-If we make the assumption that there is minimal variance in the generation interval distribution (i.e.
-$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then equation (4) can be simplified to:
-
+```{=tex}
 \begin{equation}
@@
 \end{equation}
+```
+The major advantage of this approximation is that its simplicity makes it easily understandable and verifiable.
+Anybody with basic math skills and access to the new case counts can calculate this ratio to estimate the $R_t$.
 
-We can make this estimation of $R_t$ robust to variations in the generation interval and reporting of the infection by calculating $R_t$ as the ratio of the sums of new cases within two time windows size $\tau$ days (with the center of the time periods being on days $t$ and $t-d$):
+## A Simple Ratio for Estimating Rt {.unnumbered}
 
+As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that defines the expected number of new infections created today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$).
+Therefore, to calculate $N_a$ (the total expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).
+
+```{=tex}
+\begin{equation}
+  N_a = I_{t-a}n(a)
+  (\#eq:eq2)
+\end{equation}
+```
+Then to determine the total expected number of new infections today ($I_t$), we sum $N_a$ for all values of $a$ (i.e.
+sum up the expected number of new infections created by each infected individuals based on the age of their infection).
+If we treat $a$ and $t$ instead of discrete days, we integrate over all values of $a$ to produce Equation \@ref(eq:renewal-eq).
+
+```{=tex}
+\begin{equation} 
+  I_t = \int_a^\infty{I_{t-a}n(a)}da
+  (\#eq:renewal-eq)
+\end{equation}
+```
+The generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e. $R_t$).
+
+```{=tex}
+\begin{equation} 
+  g(a) = \frac{n(a)}{R_t}
+  (\#eq:generation-interval-eq)
+\end{equation}
+```
+Thus, Equation \@ref(eq:generation-interval-eq) can be substituted into Equation \@ref(eq:renewal-eq) to produce:
+
+```{=tex}
+\begin{equation} 
+  \frac{I_t} = \int_a^\infty{I_{t-a}\frac{g(a)}{R_t}}da
+  (\#eq:sub-in-rt-1)
+\end{equation}
+```
+```{=tex}
+\begin{equation} 
+  \frac{I_t}{R_t} = \int_a^\infty{I_{t-a}g(a)}da
+  (\#eq:sub-in-rt-2)
+\end{equation}
+```
+```{=tex}
+\begin{equation} 
+  R_t = \frac{I_t}{\int{I_{t-a}g(a)}da}
+  (\#eq:sub-in-rt-3)
+\end{equation}
+```
+If we make the assumption that there is minimal variance in the generation interval distribution (i.e.
+$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then equation (4) can be simplified to:
+
+```{=tex}
+\begin{equation}
+  R_t \approx \frac{I_t}{I_{t-d}}
+  (\#eq:simple-ratio-eq)
+\end{equation}
+```
+We can make this estimation of $R_t$ robust to variations in the generation interval and reporting of new cases by calculating $R_t$ as the ratio of the sums of new cases within two time windows size $\tau$ days (with the center of the time windows being on days $t$ and $t-d$):
+
+```{=tex}
 \begin{equation} 
@@
 \end{equation}
-
+```
 A confidence interval can be generated for this ratio by carrying forward a confidence interval in the generation interval.
 
-# Materials and methods {-}
+# Materials and methods {.unnumbered}
 
-## Simulation {-}
+## Simulation {.unnumbered}
 
@@
 
-## Rt Esimation {-}
+## Rt Estimation {.unnumbered}
 
@@
 
-## COVID-19 Data {-}
+## COVID-19 Data {.unnumbered}
 
@@
 
-## Code and Data Availability {-}
+## Code and Data Availability {.unnumbered}
 
 All simulations, data processing, analysis, and graphics were all created using the R statistical software using the following packages: `r vector_to_englishList(packages)` [`r package_citation_ids`].
-All code and data used to develop this project is available at [https://github.com/aryazand/simple_rt](https://github.com/aryazand/simple_rt)
+All code and data used to develop this project is available at <https://github.com/aryazand/simple_rt>
 
-# Results {-}
+# Results {.unnumbered}
 
-## Accuracy of the simple ratio estimation of $R_t$ {-}
+## Accuracy of the simple ratio estimation of $R_t$ {.unnumbered}
 
@@
 "lockdown" period or lifting of a lockdown).
-**Figure \@ref(fig:sim-results) A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
-A full visualization of the all the simulations can be seen in (**Figure \@ref(fig:all-simulations)**).
+**Figure** \@ref(fig:sim-results) **A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
+A full visualization of the all the simulations can be seen in (**Figure** \@ref(fig:all-simulations)).
 
-Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was  `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (**Figure \@ref(fig:sim-results) B**).
-The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (**Figure \@ref(fig:sim-results) C**).
+Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (**Figure** \@ref(fig:sim-results) **B**).
+The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (**Figure** \@ref(fig:sim-results) **C**).
 This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].
@@
 
+# Descriptive Statistics
+N = simulation_parameters$initial_state_values %>% sum()
+error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
+error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
+error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
+
+# 
+rt_estimates_error_summary <- rt_estimates %>%
+  filter(time > 14) %>%
+  group_by(simulation_run) %>%
+  summarise(dm_test = list(dm.test(rt.simple_mean_error, rt.cori_mean_error)))
+
+# Plots
 panel_A <- rt_estimates %>%
@@
 
-panel_B <- rt_estimates %>%
+panel_C <- rt_estimates %>%
   filter(time > 14) %>%
@@
 
-panel_C <- rt_estimates %>%
+panel_D <- rt_estimates %>%
   filter(time > 14) %>%
@@
 
-ggarrange(panel_A, panel_B, panel_C, ncol=3, nrow=2, align = "hv", 
-          labels=c("A", "B", "C", "D", "E", "F"), common.legend = T)
+panel_D <- rt_estimates %>% 
+  pivot_longer(cols = ends_with("mean_error"), names_to = "Method", values_to = "estimation_error") %>% 
+  mutate(Method = recode(Method, rt.cori_mean_error = "cori", rt.simple_mean_error = "simple")) %>%
+  ggplot(aes(x = time, y = estimation_error)) + geom_line(aes(group = simulation_run, color = Method), alpha = 0.2) + 
+  facet_wrap(~Method, ncol = 1) +
+  custom_theme() + theme(strip.background = element_blank(), strip.text = element_blank()) + 
+  scale_color_manual(values=cbPalette) +
+  scale_fill_manual(values=cbPalette)
+
+full_figure_top = ggarrange(panel_A, panel_B, panel_C, ncol=1, nrow = 3, align = "hv", labels=c("A", "B", "C"), common.legend = T)
+full_figure_bottom = ggarrange(panel_D, labels=c("D"), common.legend = T)
+ggarrange(full_figure_top, full_figure_bottom, ncol = 2, common.legend = T)
+
 ```
@@
 Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer.
-Under these conditions, the simple ratio actually performs slightly better than than the Cori method (**Figure \@ref(fig:estimation-inaccParams)**).
+Under these conditions, the simple ratio actually performs slightly better than than the Cori method (**Figure** \@ref(fig:estimation-inaccParams)).
 
@@
 
-## Using the Simple Ratio to Estimate $R_t$ with real-world COVID-19 data {-}
+## Using the Simple Ratio to Estimate $R_t$ with real-world COVID-19 data {.unnumbered}
 
@@
 We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020].
-**Figure \@ref(fig:real-world-data) A** shows the results for the month of October 2020.
+**Figure** \@ref(fig:real-world-data) **A** shows the results for the month of October 2020.
 Visual inspection shows that the two methods of estimating $R_t$ have very good concordance.
-Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (**Figure \@ref(fig:real-world-data) B**).
+Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (**Figure** \@ref(fig:real-world-data) **B**).
 
-# Discussion {-}
+# Discussion {.unnumbered}
 
-# References {-}
+# References {.unnumbered}
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
-# Supplementary Material {-}
+# Supplementary Material {.unnumbered}
+
 \beginsupplement
 
+```{=tex}
 \newpage
 \blandscape
+```
 ```{r all-simulations, fig.width = 11, fig.height = 6, fig.cap = "All Simulations"}
@@
 ```
+
 \elandscape
@@
 \newpage
+
 ```{r estimation-inaccParams,  fig.cap = "(A) estimation with gi mean mis-specified. (B) estimation with gi sd mis-specified."}

### 2021-01-16 docs/manuscript/manuscript.Rmd
Made derivation of simple-ratio more descriptive

@@
 ```
-The generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e. $R_t$).
+Next we can substitute the generation interval function ($g(a)$) into Equation \@ref(eq:renewal-eq) by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e.
+$R_t$).
 
@@
 ```
-Thus, Equation \@ref(eq:generation-interval-eq) can be substituted into Equation \@ref(eq:renewal-eq) to produce:
-
 ```{=tex}
 \begin{equation} 
-  \frac{I_t} = \int_a^\infty{I_{t-a}\frac{g(a)}{R_t}}da
+  I_t = \int_a^\infty{I_{t-a}\frac{g(a)}{R_t}}da
   (\#eq:sub-in-rt-1)
@@
 If we make the assumption that there is minimal variance in the generation interval distribution (i.e.
-$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then equation (4) can be simplified to:
+$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \@ref(eq:renewal-eq) can be simplified to:
 

### 2021-01-16 docs/manuscript/manuscript.Rmd
improved figure 1

@@
 
-panel_C <- rt_estimates %>%
+panel_B <- rt_estimates %>%
   filter(time > 14) %>%
@@
 
-panel_D <- rt_estimates %>%
+panel_C <- rt_estimates %>%
   filter(time > 14) %>%
@@
   mutate(Method = recode(Method, rt.cori_mean_error = "cori", rt.simple_mean_error = "simple")) %>%
-  ggplot(aes(x = time, y = estimation_error)) + geom_line(aes(group = simulation_run, color = Method), alpha = 0.2) + 
-  facet_wrap(~Method, ncol = 1) +
-  custom_theme() + theme(strip.background = element_blank(), strip.text = element_blank()) + 
+  ggplot(aes(x = time, y = estimation_error)) + geom_line(aes(group = interaction(simulation_run, Method), color = Method), alpha = 0.2) + 
+  facet_wrap(~Method, ncol = 2) +
+  ylab("Estimation Error") + 
+  custom_theme() + 
+  theme(strip.background = element_blank(), 
+        strip.text = element_blank(),
+        legend.position = "none") + 
   scale_color_manual(values=cbPalette) +
-  scale_fill_manual(values=cbPalette)
+  scale_fill_manual(values=cbPalette) + 
+  ylim(-0.25, 0.5)
 
-full_figure_top = ggarrange(panel_A, panel_B, panel_C, ncol=1, nrow = 3, align = "hv", labels=c("A", "B", "C"), common.legend = T)
-full_figure_bottom = ggarrange(panel_D, labels=c("D"), common.legend = T)
-ggarrange(full_figure_top, full_figure_bottom, ncol = 2, common.legend = T)
+full_figure_top = ggarrange(panel_A, panel_B, panel_C, ncol = 3, align = "hv", labels=c("A", "B", "C"), common.legend = T)
+full_figure_bottom = ggarrange(panel_D, labels=c("D"))
+ggarrange(full_figure_top, full_figure_bottom, nrow = 2, common.legend = T)
 

### 2021-01-16 docs/manuscript/manuscript.Rmd
moved variable for proper compiling

@@
 
+```{r}
+# Descriptive Statistics
+N = simulation_parameters$initial_state_values %>% sum()
+error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
+error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
+error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
+```
+
 We assessed how accurately the simple ratio can estimate $R_t$ compared to a previously developed method by Cori et al [-@cori_new_2013], which is a popular method in the field - we will refer to this method as the "Cori method".

### 2021-01-16 docs/manuscript/manuscript.Rmd
moved variable

@@
 ```{r sim-results, fig.cap = '(ref:sim-results-cap)'}
-
-# Descriptive Statistics
-N = simulation_parameters$initial_state_values %>% sum()
-error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
-error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
-error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
-
-# 
-rt_estimates_error_summary <- rt_estimates %>%
-  filter(time > 14) %>%
-  group_by(simulation_run) %>%
-  summarise(dm_test = list(dm.test(rt.simple_mean_error, rt.cori_mean_error)))
-
 # Plots

### 2021-01-16 docs/manuscript/manuscript.Rmd
changed referencing to cref package

@@
       in_header: header.tex
+header-includes:
+  - \usepackage[labelsep=period,labelfont=bf]{caption}
+  - \usepackage{hyperref}
+  - \usepackage[capitalize, nameinlink]{cleveref}
+  - \crefdefaultlabelformat{#2\textbf{#1}#3} % <-- Only #1 in \textbf
+  - \crefname{figure}{\textbf{Figure}}{\textbf{Figures}}
+  - \Crefname{figure}{\textbf{Figure}}{\textbf{Figures}}
+  - \crefname{table}{\textbf{Table}}{\textbf{Tables}}
+  - \Crefname{table}{\textbf{Table}}{\textbf{Tables}}
 bibliography:
@@
 "lockdown" period or lifting of a lockdown).
-**Figure** \@ref(fig:sim-results) **A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
-A full visualization of the all the simulations can be seen in (**Figure** \@ref(fig:all-simulations)).
+\cref{fig:sim-results}**A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
+A full visualization of the all the simulations can be seen in (\cref{fig:all-simulations)).
 
-Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (**Figure** \@ref(fig:sim-results) **B**).
-The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (**Figure** \@ref(fig:sim-results) **C**).
+Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results)**B**).
+The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results)**C**).
 This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].
@@
 Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer.
-Under these conditions, the simple ratio actually performs slightly better than than the Cori method (**Figure** \@ref(fig:estimation-inaccParams)).
+Under these conditions, the simple ratio actually performs slightly better than than the Cori method (\cref{fig:estimation-inaccParams)).
 
@@
 We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020].
-**Figure** \@ref(fig:real-world-data) **A** shows the results for the month of October 2020.
+\cref{fig:real-world-data)**A** shows the results for the month of October 2020.
 Visual inspection shows that the two methods of estimating $R_t$ have very good concordance.
-Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (**Figure** \@ref(fig:real-world-data) **B**).
+Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (\cref{fig:real-world-data)**B**).
 

### 2021-01-16 docs/manuscript/manuscript.Rmd
improved description of how simple ratio equation is derived

@@
 
-As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that defines the expected number of new infections created today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$).
-Therefore, to calculate $N_a$ (the total expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).
+As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new infections created today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$).
+Therefore, to calculate $I_{ta}$ (the expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).
 
@@
 \begin{equation}
-  N_a = I_{t-a}n(a)
+  I_{ta} = I_{t-a}n(a)
   (\#eq:eq2)
@@
 ```
-Then to determine the total expected number of new infections today ($I_t$), we sum $N_a$ for all values of $a$ (i.e.
-sum up the expected number of new infections created by each infected individuals based on the age of their infection).
-If we treat $a$ and $t$ instead of discrete days, we integrate over all values of $a$ to produce Equation \@ref(eq:renewal-eq).
+
+Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$.
+In other words, $I_t$ can be calculated by taking all individuals infected prior to $t$, calculating the expected number of new infections each one will produce based on the age of their infection, and summing all those values together.
+If we treat $a$ as continuous variables instead of discrete days, we can integrate over all values of $a$, as shown in \cref{eq:renewal-eq-1}.
 
@@
 \begin{equation} 
-  I_t = \int_a^\infty{I_{t-a}n(a)}da
-  (\#eq:renewal-eq)
+  I_t = \int_a^\infty{I_{ta}}da
+  (\#eq:renewal-eq-1)
 \end{equation}
 ```
-Next we can substitute the generation interval function ($g(a)$) into Equation \@ref(eq:renewal-eq) by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e.
+```{=tex}
+\begin{equation} 
+  I_t = \int_a^\infty{I_{t-a}n(a)}da
+  (\#eq:renewal-eq-2)
+\end{equation}
+```
+
+Next we can substitute the generation interval function ($g(a)$) into \cref{eq:renewal-eq-2} by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e.
 $R_t$).
@@
 \begin{equation} 
-  I_t = \int_a^\infty{I_{t-a}\frac{g(a)}{R_t}}da
+  I_t = \int_a^\infty{I_{t-a}g(a)R_t}da
   (\#eq:sub-in-rt-1)
@@
 If we make the assumption that there is minimal variance in the generation interval distribution (i.e.
-$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \@ref(eq:renewal-eq) can be simplified to:
+$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \cref{eq:sub-in-rt-3} can be simplified to:
 
@@
 \cref{fig:sim-results}**A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
-A full visualization of the all the simulations can be seen in (\cref{fig:all-simulations)).
+A full visualization of the all the simulations can be seen in (\cref{fig:all-simulations}).
 
-Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results)**B**).
-The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results)**C**).
+Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results}**B**).
+The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results}**C**).
 This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].
@@
 Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer.
-Under these conditions, the simple ratio actually performs slightly better than than the Cori method (\cref{fig:estimation-inaccParams)).
+Under these conditions, the simple ratio actually performs slightly better than than the Cori method (\cref{fig:estimation-inaccParams}).
 
@@
 We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020].
-\cref{fig:real-world-data)**A** shows the results for the month of October 2020.
+\cref{fig:real-world-data}**A** shows the results for the month of October 2020.
 Visual inspection shows that the two methods of estimating $R_t$ have very good concordance.
-Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (\cref{fig:real-world-data)**B**).
+Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (\cref{fig:real-world-data}**B**).
 

### 2021-01-16 docs/manuscript/manuscript.Rmd
writing

@@
 However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
-Here, we describe how $R_t$ can be approximated by a ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (Equation 1).
-The average *generation interval* is defined as the average time from when a primary case becomes infectious to when a secondary case becomes infectious.
+Typically these methods require a description of the generation interval distribution, which is the distribution of times between when a primary case becomes infectious to when a secondary case becomes infectious.
+
+Here, we describe how $R_t$ can be approximated by a simple ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (\cref{eq:simple-ratio-eq}).
+Our aim is two-fold: First, by describing a simplified method of calculating $R_t$, it may allow others to easily and rapidly calculate $R_t$.
+Second, to assess the accuracy of the simple ratio compared to the more advanced methods for calculating $R_t$.
+This second aim is important, because this simplified ratio has been used to calculate $R_t$ in public health reports but we have not found significant literature describing the performance of the method (at least in English publications).
 
@@
 
-As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new infections created today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$).
+As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new cases produced today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$).
 Therefore, to calculate $I_{ta}$ (the expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).
@@
 ```
-
-Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$.
+Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$, as shown in \cref{eq:renewal-eq-1,eq:renewal-eq-2}.
 In other words, $I_t$ can be calculated by taking all individuals infected prior to $t$, calculating the expected number of new infections each one will produce based on the age of their infection, and summing all those values together.
-If we treat $a$ as continuous variables instead of discrete days, we can integrate over all values of $a$, as shown in \cref{eq:renewal-eq-1}.
 
@@
 \begin{equation} 
-  I_t = \int_a^\infty{I_{ta}}da
+  I_t = \sum_{a=0}^\infty{I_{ta}}da
   (\#eq:renewal-eq-1)
@@
 \begin{equation} 
-  I_t = \int_a^\infty{I_{t-a}n(a)}da
+  I_t = \sum_{a=0}^\infty{I_{t-a}n(a)}da
   (\#eq:renewal-eq-2)
@@
 ```
-
 Next we can substitute the generation interval function ($g(a)$) into \cref{eq:renewal-eq-2} by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e.
@@
 \begin{equation} 
-  I_t = \int_a^\infty{I_{t-a}g(a)R_t}da
+  I_t = \sum_{a=0}^\infty{I_{t-a}g(a)R_t}da
   (\#eq:sub-in-rt-1)
@@
 \begin{equation} 
-  \frac{I_t}{R_t} = \int_a^\infty{I_{t-a}g(a)}da
+  \frac{I_t}{R_t} = \sum_{a=0}^\infty{I_{t-a}g(a)}da
   (\#eq:sub-in-rt-2)
@@
 \begin{equation} 
-  R_t = \frac{I_t}{\int{I_{t-a}g(a)}da}
+  R_t = \frac{I_t}{\sum{I_{t-a}g(a)}da}
   (\#eq:sub-in-rt-3)
@@
 The length of the transition period was randomly selected from `r simulation_parameters$transition_lengths[1]` to `r simulation_parameters$transition_lengths[2]` days.
-Only simulations that lasted at least `r simulation_parameters$min_simLength` and never had less than `r simulation_parameters$min_newCases` cases per day were included in final analysis.
+Only simulations that lasted at least `r simulation_parameters$min_simLength` days and never had less than `r simulation_parameters$min_newCases` cases per day were included in final analysis.
 

### 2021-01-16 docs/manuscript/manuscript.Rmd
update intro

@@
 Second, to assess the accuracy of the simple ratio compared to the more advanced methods for calculating $R_t$.
-This second aim is important, because this simplified ratio has been used to calculate $R_t$ in public health reports but we have not found significant literature describing the performance of the method (at least in English publications).
+This second aim is important, because this simplified ratio has been used to calculate $R_t$ in public health reports, such as the reports created by Germany's Robert Koch Institute [@Heiden2020Schatzung], but we have not found significant literature describing the performance of the method (at least in English literature).
 
@@
 ```
-The major advantage of this approximation is that its simplicity makes it easily understandable and verifiable.
-Anybody with basic math skills and access to the new case counts can calculate this ratio to estimate the $R_t$.
-
 ## A Simple Ratio for Estimating Rt {.unnumbered}

### 2021-01-16 docs/manuscript/manuscript.Rmd
update manuscript

@@
 However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
-Typically these methods require a description of the generation interval distribution, which is the distribution of times between when a primary case becomes infectious to when a secondary case becomes infectious.
+Typically these methods require a description of the generation interval distribution, which is the distribution of times between when a primary case becomes infected to when a secondary case becomes infected.
 
@@
 ```
-## A Simple Ratio for Estimating Rt {.unnumbered}
+## Deriving the Simple Ratio for estimating $R_t$ {.unnumbered}
 
-As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new cases produced today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$).
-Therefore, to calculate $I_{ta}$ (the expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).
+A subset of previous methods for estimating $R_t$ are based on the following relationship:
 
@@
 \begin{equation} 
-  I_{ta} = I_{t-a}n(a)
-  (\#eq:eq2)
+  R_t = \frac{I_t}{\sum_{a=0}^\infty{I_{t-a}g(a)}}
+  (\#eq:original-rt)
 \end{equation}
 ```
-Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$, as shown in \cref{eq:renewal-eq-1,eq:renewal-eq-2}.
-In other words, $I_t$ can be calculated by taking all individuals infected prior to $t$, calculating the expected number of new infections each one will produce based on the age of their infection, and summing all those values together.
+$I_t$ is the number of new cases produced on day $t$; $g(a)$ is the infectiousness of a person infected $a$ days ago - $g(a)$ can be thought of as equivalent to to the generation interval distribution.
+To understand the derivation of this equation, we refer readers to [@wallinga_how_2007].
 
-```{=tex}
-\begin{equation} 
-  I_t = \sum_{a=0}^\infty{I_{ta}}da
-  (\#eq:renewal-eq-1)
-\end{equation}
-```
-```{=tex}
-\begin{equation} 
-  I_t = \sum_{a=0}^\infty{I_{t-a}n(a)}da
-  (\#eq:renewal-eq-2)
-\end{equation}
-```
-Next we can substitute the generation interval function ($g(a)$) into \cref{eq:renewal-eq-2} by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e.
-$R_t$).
-
-```{=tex}
-\begin{equation} 
-  g(a) = \frac{n(a)}{R_t}
-  (\#eq:generation-interval-eq)
-\end{equation}
-```
-```{=tex}
-\begin{equation} 
-  I_t = \sum_{a=0}^\infty{I_{t-a}g(a)R_t}da
-  (\#eq:sub-in-rt-1)
-\end{equation}
-```
-```{=tex}
-\begin{equation} 
-  \frac{I_t}{R_t} = \sum_{a=0}^\infty{I_{t-a}g(a)}da
-  (\#eq:sub-in-rt-2)
-\end{equation}
-```
-```{=tex}
-\begin{equation} 
-  R_t = \frac{I_t}{\sum{I_{t-a}g(a)}da}
-  (\#eq:sub-in-rt-3)
-\end{equation}
-```
-If we make the assumption that there is minimal variance in the generation interval distribution (i.e.
-$g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \cref{eq:sub-in-rt-3} can be simplified to:
+If we make the simplifying assumption that there is minimal variance in $g(a)$ (i.e.
+$g(a)$ is 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ is zero for all other values of $a$), then Equation \cref{eq:original-rt} can be simplified to:
 
@@
 ```
-A confidence interval can be generated for this ratio by carrying forward a confidence interval in the generation interval.
+A confidence interval can be generated for this ratio by carrying forward a confidence interval of the mean generation interval.
 
@@
 
-## Rt Estimation {.unnumbered}
+## $R_t$ Estimation {.unnumbered}
 
@@
 
-Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results}**B**).
+Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (95% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results}**B**).
 The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results}**C**).
 This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].
+When $R_t$ is relatively stable, the simple method may over-estimate $R_t$; however during periods of rapid change in $R_t$ (e.g.
+going in and out of lockdown), the simple method may either over- or under-estimate $R_t$ (\cref{fig:sim-results}**D**).
 
-(ref:sim-results-cap) \textbf{The Simple Ratio approximates the effective reproduction in simualted epidemics.} One hundred epidemics were simulated using an SEIR model and the simple ratio and Cori method were used to estimate the known $R_t$ over time for each simulated epidemic.
-\textbf{(A)} The true $R_t$ (black line) and estimated $R_t$ (colored lines) over time for one of simulated epidemics.
+(ref:sim-results-cap): \textbf{The Simple Ratio approximates the effective reproduction in simualted epidemics.} One hundred epidemics were simulated using an SEIR model and the simple ratio and Cori method were used to estimate the known \$R_t\$ over time for each simulated epidemic.
+\textbf{(A)} The true \$R_t\$ (black line) and estimated \$R_t\$ (colored lines) over time for one of simulated epidemics.
 \textbf{(B)} The distribution of the absolute errors of both estimation methods across all simulated epidemics.
-\textbf{(C)} The estimation error of the simple ratio method is associated with size of true $R_t$
+\textbf{(C)} The estimation error of the simple ratio method is associated with size of true \$R_t\$.
+\textbf{(D)} The estimation error of both methods over time for the simulated epidemics.
+Each line represents the error in \$R_t\$ estimation for a given simulation.
+Note that for both methods, a substantial proportion of error occurs around days 60 and 130 when there was is rapid change in \$R_t\$
 
-```{r sim-results, fig.cap = '(ref:sim-results-cap)'}
+```{r sim-results, fig.cap = '(ref: sim-results-cap)'}
+
 # Plots
@@
 
-```{r estimation-inaccParams,  fig.cap = "(A) estimation with gi mean mis-specified. (B) estimation with gi sd mis-specified."}
+```{r estimation-inaccParams,  fig.cap = "\\textbf{$R_t$ estimation error when the mean or standard deviation of the generation interval is mis-specified}.
+Similar to \\textbf{Figure 1}, this figure demonstrate the distribution of absolute error when estimating $R_t$ for 100 simulated epidemics.
+\\textbf{(A)} Distribution of errors when the mean of the generation interval is mis-specified as 2 days too long.
+The average \\textbf{(B)} Distribution of errors when the standard deviation is mis-specified as 3 days too long."}
 

### 2021-01-21 docs/manuscript/manuscript.Rmd
update manuscript

@@
   ABSTRACT
-acknowledgements: |
-  Acknowledgements
 output:
-  bookdown::pdf_book:
-    base_format: rticles::oup_article
+  bookdown::word_document2:
     number_sections: no
@@
       in_header: header.tex
-  bookdown::word_document2:
+  bookdown::pdf_book:
+    base_format: rticles::oup_article
     number_sections: no
@@
 face mask usage, physical distancing, and travel restrictions).
-If the $R_t$ is above 1, it indicates that the epidemic is growing exponentially.
-If, by contrast, the $R_t$ drop below 1, it indicates that the number of new cases is decreasing.
+While the $R_t$ is a retrospective measure and does not predict future behavior of an epidemic, observing the $R_t$ consistently above 1, indicates that the epidemic is growing exponentially.
+If, by contrast, the $R_t$ is consistently below 1, it indicates that epidemic is shrinking.
 Furthermore, the $R_t$ can vary between geographic locations, as social behaviors and environmental factors vary between these locations.
 
-From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region.
+As the COVID-19 pandemic has demonstrated, communication to the public and political leadership regarding infection rates is critical to management of the pandemic.
+The $R_t$ can be used to communicate whether an epidemic is growing or shrinking in a region.
 However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007].
-Typically these methods require a description of the generation interval distribution, which is the distribution of times between when a primary case becomes infected to when a secondary case becomes infected.
 
 Here, we describe how $R_t$ can be approximated by a simple ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (\cref{eq:simple-ratio-eq}).
-Our aim is two-fold: First, by describing a simplified method of calculating $R_t$, it may allow others to easily and rapidly calculate $R_t$.
-Second, to assess the accuracy of the simple ratio compared to the more advanced methods for calculating $R_t$.
-This second aim is important, because this simplified ratio has been used to calculate $R_t$ in public health reports, such as the reports created by Germany's Robert Koch Institute [@Heiden2020Schatzung], but we have not found significant literature describing the performance of the method (at least in English literature).
+The *generational interval* being the time between when a primary case infected to the when they infect a secondary case.
+
+Our aim is two-fold: First, by describing a simplified method of calculating $R_t$, we aim to allow others to easily and rapidly calculate $R_t$.
+Second, we aim to assess the accuracy of the simple ratio compared to the more advanced methods for estimating the $R_t$.
+This second aim is important, because this simplified ratio has been used by public health bodies to report the $R_t$ during the COVID-19 pandemic - e.g.
+reports created by Germany's Robert Koch Institute [@Heiden2020Schatzung] - but we have not found significant literature describing the performance of the method (at least in published English literature).
 
@@
 
-A subset of previous methods for estimating $R_t$ are based on the following relationship:
+Among the previously developed methods estimating $R_t$, it common to base the estimation on the following relationship:
 
@@
 ```
-$I_t$ is the number of new cases produced on day $t$; $g(a)$ is the infectiousness of a person infected $a$ days ago - $g(a)$ can be thought of as equivalent to to the generation interval distribution.
-To understand the derivation of this equation, we refer readers to [@wallinga_how_2007].
+$I_t$ is the number of new cases produced on day $t$; $g(a)$ is the probability density function describing the probability that a person infected $a$ days ago will cause a new case today (at time $t$) - $g(a)$ can be thought of as equivalent to to the generation interval distribution.
+Intuitively, this equations can be understood in the following way.
+Consider a scenario where for every day prior to today, only one person was infected.
+In such a scenario, the denominator of \cref{eq:original-rt} would equal 1 (since it would be the sum of all values of probability density function) and the $R_t$ would equal $I_t$.
+In this framework $R_t$ is equal to the number of new infections created today by a cohort of people that span the entire lifetime of an infection, which is mathematically equivalent to the traditional framework of $R_t$ as the total number of new cases generated by one infected person across the entire lifetime of that person's infection.
+Now consider the scenario where multiple people were infected each day prior to today.
+In such a scenario, each additional infected person would contribute an additional probability of new cases and increase the value of $I_t$; therefore to estimate $R_t$, $I_t$ must be scaled down by the additional probability each previously infected person contributes - hence the denominator in \cref{eq:original-rt}.
+To understand the mathematical derivation of this equation, we refer readers to the article by Wallinga and Lipsitch [-@wallinga_how_2007].
 
@@
 N = simulation_parameters$initial_state_values %>% sum()
-error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
-error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
-error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
+error_summary = rt_estimates %>% filter(time > 49) %>% 
+  select(matches("rt.+_mean_error")) %>% 
+  rename_with(~ gsub("rt\\.(.+)_mean_error", "\\1", .x)) %>%
+  transmute_all(abs) %>% 
+  summarise_all(quantile, probs = c(0.025, 0.5, 0.975), na.rm = T) %>% 
+  transmute_all(signif, digits = 2)
 ```
 
-We assessed how accurately the simple ratio can estimate $R_t$ compared to a previously developed method by Cori et al [-@cori_new_2013], which is a popular method in the field - we will refer to this method as the "Cori method".
+We assessed how accurately the simple ratio can estimate $R_t$ compared to a standard method developed by Cori et al [-@cori_new_2013] - which we will refer to as the "Cori method".
 $R_t$ calculation with either the simple ratio and Cori method requires knowing the mean generation interval of the infection; however, the Cori method also requires knowing the standard deviation of the generation interval or having an empiric distribution of the generation interval.
@@
 \cref{fig:sim-results}**A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic.
-A full visualization of the all the simulations can be seen in (\cref{fig:all-simulations}).
+A full visualization of all the simulations can be seen in (\cref{fig:all-simulations}).
 
-Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (95% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results}**B**).
-The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results}**C**).
+Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_summary$simple['50%']` (95% inter-percentile range of `r error_summary$simple['2.5%']` to `r error_summary$simple['97.5%']`) (\cref{fig:sim-results}**B**).
+By comparison, the median estimation error of the Cori method was `r error_summary$cori['50%']` (95% inter-percentile range of `r error_summary$cori['2.5%']` to `r error_summary$cori['97.5%']`).
+
+We identified two factors contributing to error by the simple ratio.
+First, the simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results}**C**).
 This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].
+Second, the simple method (like the Cori method) is more prone to error when there is a rapid change in $R_t$ - i.e.
+going in and out of lockdown (\cref{fig:sim-results}**D**).
 When $R_t$ is relatively stable, the simple method may over-estimate $R_t$; however during periods of rapid change in $R_t$ (e.g.
-going in and out of lockdown), the simple method may either over- or under-estimate $R_t$ (\cref{fig:sim-results}**D**).
+going in and out of lockdown), the simple method may either over- or under-estimate $R_t$.
+Overall, while the Cori method performs better, the simple method provides an estimation within a reasonable margin of error.
 
-(ref:sim-results-cap): \textbf{The Simple Ratio approximates the effective reproduction in simualted epidemics.} One hundred epidemics were simulated using an SEIR model and the simple ratio and Cori method were used to estimate the known \$R_t\$ over time for each simulated epidemic.
-\textbf{(A)} The true \$R_t\$ (black line) and estimated \$R_t\$ (colored lines) over time for one of simulated epidemics.
+(ref:sim-results-cap) \textbf{The Simple Ratio approximates the effective reproduction in simualted epidemics.} One hundred epidemics were simulated using an SEIR model and the simple ratio and Cori method were used to estimate the known $R_t$ over time for each simulated epidemic.
+\textbf{(A)} The true $R_t$ (black line) and estimated $R_t$ (colored lines) over time for one of simulated epidemics.
 \textbf{(B)} The distribution of the absolute errors of both estimation methods across all simulated epidemics.
-\textbf{(C)} The estimation error of the simple ratio method is associated with size of true \$R_t\$.
+The median absolute error for the Cori and simple method were `r error_summary$cori['50%']` (95% IPR `r error_summary$cori['2.5%']` to `r error_summary$cori['97.5%']`) and `r error_summary$simple['50%']` (95% IPR `r error_summary$simple['2.5%']` to `r error_summary$simple['97.5%']`), respectively.
+Note only error after the 50th day was considered because error early in the pandemic was inordinately large for both methods, as shown in Panel D.
+\textbf{(C)} The estimation error of the simple ratio method is associated with size of true $R_t$.
 \textbf{(D)} The estimation error of both methods over time for the simulated epidemics.
-Each line represents the error in \$R_t\$ estimation for a given simulation.
-Note that for both methods, a substantial proportion of error occurs around days 60 and 130 when there was is rapid change in \$R_t\$
+Each line represents the error in $R_t$ estimation for a given simulation.
+Note that for both methods, a substantial proportion of error occurs around days 60 and 130 when there was is rapid change in $R_t$.
+IPR = "inter-percentile range".
 
-```{r sim-results, fig.cap = '(ref: sim-results-cap)'}
+```{r sim-results, fig.cap = '(ref:sim-results-cap)'}
 
@@
 panel_B <- rt_estimates %>%
-  filter(time > 14) %>%
+  filter(time > 49) %>%
   create_rt_estimation_error_boxplot()
@@
 panel_C <- rt_estimates %>%
-  filter(time > 14) %>%
+  filter(time > 49) %>%
   create_rt_estimation_error_by_rt()
@@
 
-Often however, the the mean and standard deviation of the generation interval is typically not known exactly.
+Often the the mean and standard deviation of the generation interval is not known exactly.
 For instance, for the COVID-19 the mean generation interval estimates, with 95% credible intervals, include 3.95 (3.01 - 4.91) and 5.20 (3.78 - 6.78); and estimates of the standard deviation of the generation interval include 1.51 (0.74 - 2.97) and 1.72 (0.91 - 3.93) [@ganyani_estimating_2020].
 Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer.
-Under these conditions, the simple ratio actually performs slightly better than than the Cori method (\cref{fig:estimation-inaccParams}).
+Under these conditions, the simple ratio actually performs almost identically to the Cori method (\cref{fig:estimation-inaccParams}).
 
-```{r real-world-data, fig.cap = "\\textbf{The Simple Ratio and Cori Method estimate similar Rt values for the COVID-19 pandemic}.
-\\textbf{(A)} Rt esimations for four sample nations \\textbf{(B)} Histogram displaying the absolute difference in Rt esimation across for each day of the pandemic (Apr 01 to Dec 12) for all 233 individual nations."}
+(ref:real-world-data-cap) \textbf{The Simple Ratio and Cori Method estimate similar Rt values for the COVID-19 pandemic}.
+\textbf{(A)} Rt estimations for four sample nations \textbf{(B)} Violin plot displaying the absolute difference in Rt estimation across for each day of the pandemic (Apr 01 to Dec 12) for all 233 individual nations.
+Median absolute difference in estimations was `r estimation_diff_median` (95% IPR `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`).
+IPR = "inter-percentile range".
+
+```{r real-world-data, fig.cap = "(ref:real-world-data-cap)"}
 
@@
     geom_hline(yintercept = 1, linetype = 2) + 
-    facet_wrap(~country) + 
+    facet_wrap(~country, scales = "free") + 
     ylab("Rt") + 
@@
 panel_B <- national_data_filtered %>%
-  ggplot() + 
-    geom_histogram(aes(estimation_diff)) + scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) + 
-    labs(x = "absolute difference in Rt estimation") +
+  ggplot(aes("simple vs cori", estimation_diff)) + 
+    geom_violin() +
+    geom_boxplot(width = 0.1, outlier.alpha = 0) +
+    scale_y_log10(labels = scales::label_number(accuracy = 0.0001)) + 
+    labs(y = "absolute difference in Rt estimation", x = "") +
     custom_theme()
 
-ggarrange(panel_A, ggarrange(panel_B, ncol=2), nrow = 2, heights = c(2,1), labels = c("A","B"))
+ggarrange(panel_A, panel_B, ncol = 2, widths = c(2,1), labels = c("A","B"), common.legend = T)
 ```
@@
 
+\newpage
+
 # Supplementary Material {.unnumbered}
@@
 ```
+```{r load-supp-data}
+rt_estimates_meanerr <- readd("rt_estimates_meanerr")
+rt_estimates_sderr <- readd("rt_estimates_sderr")
+
+supp_error_summary.meanerr = rt_estimates_meanerr %>% filter(time > 49) %>% 
+  select(matches("rt.+_mean_error")) %>% 
+  rename_with(~ gsub("rt\\.(.+)_mean_error", "\\1", .x)) %>%
+  transmute_all(abs) %>% 
+  summarise_all(quantile, probs = c(0.025, 0.5, 0.975), na.rm = T) %>% 
+  transmute_all(signif, digits = 2) %>%
+  transmute_all(format, scentific = F)
+
+supp_error_summary.sderr = rt_estimates_sderr %>% filter(time > 49) %>% 
+  select(matches("rt.+_mean_error")) %>% 
+  rename_with(~ gsub("rt\\.(.+)_mean_error", "\\1", .x)) %>%
+  transmute_all(abs) %>% 
+  summarise_all(quantile, probs = c(0.025, 0.5, 0.975), na.rm = T) %>% 
+  transmute_all(signif, digits = 2) %>%
+  transmute_all(format, scentific = F)
+```
+
 ```{r all-simulations, fig.width = 11, fig.height = 6, fig.cap = "All Simulations"}
@@
 
-```{r estimation-inaccParams,  fig.cap = "\\textbf{$R_t$ estimation error when the mean or standard deviation of the generation interval is mis-specified}.
-Similar to \\textbf{Figure 1}, this figure demonstrate the distribution of absolute error when estimating $R_t$ for 100 simulated epidemics.
-\\textbf{(A)} Distribution of errors when the mean of the generation interval is mis-specified as 2 days too long.
-The average \\textbf{(B)} Distribution of errors when the standard deviation is mis-specified as 3 days too long."}
+(ref:estimation-inaccParams-cap) \textbf{$R_t$ estimation error when the mean or standard deviation of the generation interval is mis-specified}.
+Similar to \textbf{Figure 1}, this figure demonstrate the distribution of absolute error when estimating $R_t$ for 100 simulated epidemics.
+\textbf{(A)} Distribution of errors when the mean of the generation interval is mis-specified as 2 days too long.The median absolute error for the Cori and simple method were `r supp_error_summary.meanerr$cori['50%']` (95% IPR `r supp_error_summary.meanerr$cori['2.5%']` to `r supp_error_summary.meanerr$cori['97.5%']`) and `r supp_error_summary.meanerr$simple['50%']` (95% IPR `r supp_error_summary.meanerr$simple['2.5%']` to `r supp_error_summary.meanerr$simple['97.5%']`), respectively.
+The average \textbf{(B)} Distribution of errors when the standard deviation is mis-specified as 3 days too long.
+The median absolute error for the Cori and simple method were `r supp_error_summary.sderr$cori['50%']` (95% IPR `r supp_error_summary.sderr$cori['2.5%']` to `r supp_error_summary.sderr$cori['97.5%']`) and `r supp_error_summary.sderr$simple['50%']` (95% IPR `r supp_error_summary.sderr$simple['2.5%']` to `r supp_error_summary.sderr$simple['97.5%']`), respectively.
 
-rt_estimates_meanerr <- readd("rt_estimates_meanerr")
-rt_estimates_sderr <- readd("rt_estimates_sderr")
-
+```{r estimation-inaccParams,  fig.cap = "(ref:estimation-inaccParams-cap)"}
 panel_A <- rt_estimates_meanerr %>% 
+  filter(time > 49) %>%
   create_rt_estimation_error_boxplot()
@@
 panel_B <- rt_estimates_sderr %>% 
+  filter(time > 49) %>%
   create_rt_estimation_error_boxplot()

### 2021-01-21 code/create_figures.R
make figure axis log scale

@@ -25,13 +25,13 @@ create_rt_estimation_error_boxplot <- function(data) {
     pivot_longer(., cols = starts_with("rt."), names_to = c("Method", "parameter"),
                  names_pattern = "rt.([[:alpha:]]+)_([[:print:]]+)", values_to = "value") %>%
     pivot_wider(names_from = "parameter", values_from = "value") %>%
-    mutate(error = abs(mean_error)) %>%
+    mutate(error = log10(abs(mean_error))) %>%
     ggplot(., aes(Method, error)) +
     geom_violin(aes(fill = Method)) +
     geom_boxplot(width = 0.1, outlier.alpha = 0.5) +
-    ylab("log10(Estimation Error)") +
+    ylab("Absolute Estimation Error (log10)") +
     custom_theme() +
-    scale_y_log10(labels = scales::label_number(accuracy = 0.00001)) +
+    #scale_y_log10(labels = scales::label_number(accuracy = 0.00001)) +
     scale_color_manual(values=cbPalette) +
     scale_fill_manual(values=cbPalette)
 }

### 2021-01-21 code/functions.R
as.tibble to as_tibble

@@ -124,7 +124,7 @@ model_pandemic <- function(initial_state_values, mean_gi, gamma_val, r0, period_
     mutate(new_cases = c(NA, diff(total_infected)))
 
 
-  output = as.tibble(output)
+  output = as_tibble(output)
   attr(output, "metadata") = c(mean_gi = mean_gi, gamma_val = gamma_val)
 
   return(output)

### 2021-01-21 code/plan.R
update error parameters in analysis plan

@@ -34,7 +34,7 @@ plan <- drake_plan(
   rt_estimates_sderr = map(epidemic_simulations, function(i) estimates_rt(i,
                            mean_gi = attributes(i)$metadata["mean_gi"],
                            gamma_val = attributes(i)$metadata["gamma_val"], tau = 7,
-                           error.mean = 0, error.sd = 2)) %>% bind_rows(., .id = "simulation_run"),
+                           error.mean = 0, error.sd = 3)) %>% bind_rows(., .id = "simulation_run"),
 
   # Real-world data
   COVID19_params = list(mean_gi = 4.83, sd_gi = 1.72),
